```{r knitsetup, echo=FALSE, results='hide', warning=FALSE, message=FALSE, cache=FALSE}
opts_knit$set(base.dir='./', fig.path='', out.format='md')
opts_chunk$set(prompt=FALSE, comment='', results='markup')
# See yihui.name/knitr/options for more Knitr options.
##### Put other setup R code here

library(knitcitations)
options("citation_format" = "compatibility")
bib <- read.bibtex("../references.bib")

# end setup chunk
```
# Whittaker-based gap-filling

As outlined in an earlier chapter, NDVI3g.v1 (NDVI3g.v0) pixels with a flag value other than 0 (1 and 2) likely suffer from cloud/snow interference and hence originate from either spline interpolation or the average seasonal profile. These values are not considered reliable `r citep(list(bib["Detsch2016a"], bib["Fensholt2012"]))` and should hence be discarded before any further analysis, e.g. by specifying 'keep' in `rasterizeGimms`.

Among the variety of readily available approaches to fill the thus introduced gaps in vegetation indices time series -- see `r citet(bib["Liu2017"])` for an up-to-date overview of possible approaches -- the following code snippet utilizes the so-called 'Modified Whittaker Smoother' `r citep(bib["Atzberger2011"])`. The algorithm is provided through `whittaker.raster` which is part of the **MODIS** package and is hereinafter operated with an annual lambda of 500. See the function documentation for further details.

```{r mannKendall, eval = FALSE}
### download data -----

## number of cores for parallel processing
cores <- parallel::detectCores() - 1

## download entire gimms ndvi3g collection in parallel
nc4 <- downloadGimms(x = as.Date("1982-01-01"), cores = cores)


### rasterize images including quality control -----

## reference extent
library(rworldmap)
deu <- subset(countriesCoarse, ADMIN == "Germany")

## rasterize
tfs <- rasterizeGimms(nc4, ext = deu, keep = 0, cores = cores,
                      filename = gsub(".nc4", ".tif", nc4),
                      format = "GTiff", overwrite = TRUE)


### perform gap-filling -----

library(MODIS)

## time stamps
tms <- monthlyIndices(nc4, timestamp = TRUE)
otm <- orgTime(tms)

## apply whittaker smoother
wht <- whittaker.raster(tfs, timeInfo = otm, lambda = 500L, outDirPath = "wht")
```

<center>
  <img src="http://imgur.com/xhCvlIB" alt="whittaker" style="width: 600px;"/><br>
  <b>Figure 3.</b> Raw NDVI3g.v1 (gray; 2000 to 2009) and Whittaker-smoothed curve (black) for 12 random pixels in Germany.
</center>
