```{r knitsetup, echo=FALSE, results='hide', warning=FALSE, message=FALSE, cache=FALSE}
opts_knit$set(base.dir='./', fig.path='', out.format='md')
opts_chunk$set(prompt=FALSE, comment='', results='markup')
# See yihui.name/knitr/options for more Knitr options.
##### Put other setup R code here

library(knitcitations)
options("citation_format" = "compatibility")
bib <- read.bibtex("../references.bib")

# end setup chunk
```

# Quality control

Using the default settings, `rasterizeGimms` creates proper `Raster*` objects from the ENVI binary images. Associated with each half-monthly NDVI<sub>3g</sub> dataset is a quality flag layer indicating the data reliability on a pixel basis. The flag values are as follows (taken from <http://ecocast.arc.nasa.gov/data/pub/gimms/3g.v0/00READMEgeo.txt>):
  
  Flag value      | <div align="left">Description</div>
    --------------- | -----------    
    1 and 2 | good value
  3 | value derived from spline interpolation
  4 | value derived from spline interpolation, possibly snow
  5 | value derived from average seasonal profile
  6 | value derived from average seasonal profile, possibly snow
  7 | missing value
  
Quality control can easily be carried out by invoking the function `qualityControl` which works both on binary files and already rasterized images. With regard to memory usage, however, we strongly advice to create NDVI and flag layers separately via `rasterizeGimms` (with `filename` provided) and use these as input for `qualityControl`.   

```{r available_files, echo = FALSE}
gimms_files <- rearrangeFiles(full.names = TRUE, dsn = "/media/permanent/programming/r/gimms/data/", 
                              pattern = "^geo81.*VI3g$")
```

```{r quality_control}
ndvi <- rasterizeGimms(gimms_files[1])
flag <- rasterizeGimms(gimms_files[1], flag = TRUE)

qc <- qualityControl(ndvi, flag)
```

```{r visualize, echo = FALSE}
library(RColorBrewer)
library(grid)
library(lattice)

p_ndvi <- spplot(ndvi, col.regions = colorRampPalette(brewer.pal(11, "BrBG")), 
                 at = seq(-0.2, 1, 0.01), scales = list(draw = TRUE), 
                 colorkey = list(space = "top", width = .6, height = .6))
p_flag <- spplot(flag, col.regions = terrain.colors(7), at = .5:7.5, 
                 scales = list(draw = TRUE, 
                               y = list(col = "transparent")), 
                 colorkey = list(space = "top", width = .6, height = .6))
p_qc <- spplot(qc, col.regions = colorRampPalette(brewer.pal(11, "BrBG")), 
                 at = seq(-0.2, 1, 0.01), scales = list(draw = TRUE), 
               colorkey = list(space = "bottom", width = .6, height = .6))

## ndvi
grid.newpage()
vp_ndvi <- viewport(x = 0, y = .45, width = .55, height = .55, 
                    just = c("left", "bottom"))
pushViewport(vp_ndvi)
print(p_ndvi, newpage = FALSE)
downViewport(trellis.vpname("figure"))
grid.text("a) Raw NDVI", x = .5, y = 1.55, gp = gpar(cex = .8))

## flag
upViewport(0)
vp_flag <- viewport(x = .45, y = .45, width = .55, height = .55, 
                    just = c("left", "bottom"))
pushViewport(vp_flag)
print(p_flag, newpage = FALSE)
downViewport(trellis.vpname("figure"))
grid.text("b) Flags", x = .5, y = 1.55, gp = gpar(cex = .8))

## qc
upViewport(0)
vp_qc <- viewport(x = 0, y = 0.01, width = 1, height = .625, 
                  just = c("left", "bottom"))
pushViewport(vp_qc)
print(p_qc, newpage = FALSE)
downViewport(trellis.vpname("figure"))
grid.text("c) Quality-controlled NDVI", x = .5, y = -.29, gp = gpar(cex = .8))
```