```{r knitsetup, echo=FALSE, results='hide', warning=FALSE, message=FALSE, cache=FALSE}
opts_knit$set(base.dir='./', fig.path='', out.format='md')
opts_chunk$set(prompt=FALSE, comment='', results='markup')
# See yihui.name/knitr/options for more Knitr options.
##### Put other setup R code here
library(gimms)

# end setup chunk
```
# What's the story?

### Background
I've been collecting functions related to the download and processing of AVHRR GIMMS data for quite some time and with the most recent update to NDVI<sub>3g</sub> `r citep("10.3390/rs6086929")`, it was probably a good time to stuff the most fundamental work steps into a proper R package. In this context, 'most fundamental' refers to certain operations which I tended to repeat over and over again in the context of GIMMS data processing, including

* list all bi-monthly files available online at [NASA ECOCAST](http://ecocast.arc.nasa.gov/data/pub/gimms/3g.v0/),
* download selected (if not all) files, 
* re-arrange downloaded files according to date, 
* transform ENVI binary data to proper **raster** format `r citep(citation("raster"))` and 
* aggregate bi-monthly datasets to monthly maximum value composites (MVC).

In the following, you will find a short introduction of what I came up with so far. Feel free to contact me directly via <florian.detsch(at)staff.uni-marburg.de> or raise issues and provide (constructive) criticism on [GitHub](https://github.com/environmentalinformatics-marburg/gimms). Any suggestions on how to improve the **gimms** package are highly appreciated!

### How to install
The **gimms** package (currently version 0.2.0) is now officially on CRAN and can be install directly via 

```{r install, message = FALSE, eval = FALSE} 
## install latest stable package release
install.packages("gimms")
```

The development version including latest functionality updates, bug-fixes etc. can be installed directly from GitHub via **devtools** `r citep(citation("devtools"))`. Make sure to check out the [introductory page](https://github.com/environmentalinformatics-marburg/gimms) for the latest news and updates!

```{r install_dev, message = FALSE, eval = FALSE}
## install package development version
library(devtools)
install_github("environmentalinformatics-marburg/gimms", ref = "develop")
```

```{r references, echo = FALSE, eval = FALSE}
## update bibliography
knitcitations::write.bibtex(file = "references.bib", append = FALSE)
```